---
title: ' Data processing'
author: "Zia Ahmed, PhD, University at Buffalo"
date: "April 5, 2018"
output:
  html_document: default
  word_document: default
---

#### Load R packages 

```{r message=F, warning=F}
library(rgdal)  # spatial data processing
library(raster) # raster processing
library(plyr)   # data manipulation 
library(dplyr)  # data manipulation 
library(RStoolbox) # ploting spatial data 
library(RColorBrewer)
library(ggplot2)
library(sp)
```

#### Set working directory

```{r}
setwd("F:\\My_GitHub\\DNN_H20_R")
dsn<-("F:\\My_GitHub\\DNN_H20_R\\Data")
```

#### Import training polygon layer 

```{r}
poly <- readOGR(dsn=dsn, "Training_layer") 
ID<-read.csv("Data\\Landuse_ID.csv", header=T)
ID
```

#### Convert polygon to raster
#### Raster extent (we will use any band of Sentinel-2 to set raster extent, we use B2 for define extent)

```{r}
b2<-raster("Data\\B2.tif") # 
crs(b2) <- "+proj=utm +zone=17N +ellps=WGS84 +datum=WGS84 +units=m +no_defs" 
extent=extent(b2)
```

#### Convert to raster (2.5 m grid size)

```{r}
r <- raster(extent, resolution=2.5,
            crs = '+proj=utm +zone=17N +ellps=WGS84 +datum=WGS84 +units=m +no_defs ')
extent(r) <- extent(poly)
rp <- rasterize(poly, r, 'Class_ID')
plot(rp)
```

#### Convert raster to data.frame and rename colum to "layer"" to Class_ID

```{r}
rp.df <- as.data.frame(rasterToPoints(rp))
colnames(rp.df)[3] <- 'Class_ID'
```

#### Create a Spatial point Data frame

```{r}
xy <- rp.df[,c(1,2)]
point.SPDF <- SpatialPointsDataFrame(coords = xy,
                                data=rp.df,
                               proj4string = CRS("+proj=utm +zone=17N +ellps=WGS84 +datum=WGS84 +units=m +no_defs"))
```

#### Load multi-band raster 

```{r}
multi=stack("Data//multi_bands.tif")
# rename the bands
names(multi) <- c("B2", "B3","B4","B5","B6","B7","B8","B8A","B11","B12")
```

#### Plot map

```{r}

# Natural Color
p1<-ggRGB(multi, r=3, g=2, b=1, stretch = "lin")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  ggtitle("Natural Color\n(R= Red, G= Green, B= Blue)")
# False Color image
p2<-ggRGB(multi, r=7, g=3, b=2, stretch = "lin")+
  theme(axis.title.x=element_blank(),
        axis.text.x=element_blank(),
        axis.ticks.x=element_blank(),
        axis.title.y=element_blank(),
        axis.text.y=element_blank(),
        axis.ticks.y=element_blank())+
  ggtitle("False Color Infrared\n(R= NIR, G= Red,  B=Green)")
source("multi_plot_function.r") # We need this function to plot multiple figures in one page
#(http://www.cookbook-r.com/Graphs/Multiple_graphs_on_one_page_(ggplot2)/)
multiplot(p1, p2, cols=2)
```

####  Save as a tiff file

```{r}
windows(width=6, height=4)
tiff( file="FIGURE_Natural_False.tif", 
      width=6, 
      height=4,
      units = "in", 
      pointsize = 14, 
      res=1600, bg = "white",
      restoreConsole = TRUE,
      compression =  "lzw")
par(mar=c(5,5,2,1), oma=c(0,0,0,0))
multiplot(p1, p2, cols=2)
dev.off()

```

#### Extract raster values to point file

```{r}
point.df <- extract(multi, point.SPDF, df=TRUE, method='simple')
```

#### Combine with data frame

```{r}
point.mf<-cbind(rp.df,point.df)
head(point.mf)
```

#### Keep values belong only to 5 classes

```{r}
point.train<-point.mf %>% 
  select(x,y,Class_ID, B2, B3, B4, B5, B6,B7,B8,B8A,B11,B12) %>% 
  filter(Class_ID >0) 
```

#### Add class ID and save as a CSV file 

```{r}
point<-join(point.train, ID, by="Class_ID", type="inner")
write.csv(point, "point_data.csv", row.names=F)
```

### Convert raster stack to CSV file
#### First creat XY data-frame, will use Band B2

```{r}
grid.point <- data.frame(rasterToPoints(b2))
# Remove B2 column
grid.point$B2<-NULL
```

#### Convert to sptial point data frame and define projection

```{r}
coordinates(grid.point) <- ~x + y
projection(grid.point) <- CRS("+proj=utm +zone=17N +ellps=WGS84 +datum=WGS84 +units=m +no_defs")
```

#### Extract all bands values to grird.point

```{r}
df.grid<- extract(multi, grid.point, df=TRUE, method='simple')
```

#### Combine with grid.df (we need to add xy coordinated for mapping) and write as a CSV file

```{r}
grid<-cbind(as.data.frame(grid.point),df.grid)
write.csv(grid, "grid_data.csv", row.names=F)
head(grid)
```


